\section{Bitcoin and Blockchain scalability}
\subsection{Introduction}
As a consequence of the increasing adoption of Blockchain-based cryptocurrencies,
their ability to scale has raised concerns and has received a lot of attention
in the last few years. In particular, the key concerns are:
\begin{itemize}
  \item[-] can cryptocurrencies based on decentralized blockchains be scaled up
  to match the performance of a mainstream payment processor?
  \item[-] what does it take to get there?
\end{itemize}

\paragraph{Bitcoin current performance} As reference, Bitcoin today requires
around 10 minutes to confirm a transaction (a new block is mined every $\sim10$
minutes) and achieves a maximum throughput of 7 transaction/sec
\cite{wikipedia_scalability_2018}. Since the transactions are confirmed only
after the block they belong to is created and added to the blockchain, the
maximum throughput of Bitcoin is effectively capped at maximum block size
divided by block interval. In comparison, a payment processor such as Visa
credit card processes 2000 transactions/sec on average, with a peak rate of
56,000 transactions/sec \cite{wikipedia_scalability_2018}.

\paragraph{Bitcoin reparametrization} A solution for increasing Bitcoin
throughput could therefore changing the block interval time (currently 10
minutes) and increasing upper limit of the block size, which currently is 1MB.
In the last few years there's been a debate about this topic which split the
community. People in favour for increasing the size claim that increasing it
would allow Bitcoin to easily reach VISA (and analog payment systems) numbers,
while the opposing ones claim that this would damage decentralization because
blocks of big size require a lot of computational power for being mined and this
increases the costs of participation, centralizing the miners in a few powerful
nodes. Their proposed solutions, therefore, consist of spending effort for
optimizing the use of the current block space available and offloading certain
processing to off-chain networks (off-chain solutions).

As discussed in \cite{croman-scaling-blockchain}, since scalability is not a
singular metric and it includes various performance and security metrics,
reparametrization can achieve only limited benefits considering the network
performance given by Bitcoin’s current peer-to-peer network protocol and the
willing to maintain its current degree of decentralization. However, it is still
an open question whether reparametrization alone can address the growth of
Bitcoin to the same order of magnitude of systems like the previously mentioned
VISA. Following the considerations discussedssed in
\cite{croman-scaling-blockchain}, the next section will explore the
reparametrization limitations which show that likely the scaling problem of
Bitcoin (and, more in general, of Blockchain systems) cannot be faced with
reparametrization alone.

\subsection{Reparametrization limitations}

\subsubsection{Current Bitcoin metrics}
\paragraph{Maximum throughput}  Today Bitcoin’s maximum throughput is 7
transactions/sec \cite{wikipedia_scalability_2018}. This number is limited by
the maximum block size and the interblock time.

\paragraph{Maximum and average block size} The maximum block size is 1 MB, while
the average size today is around 800 KB \cite{current-block-size}. In the years
2014-2015, when the measurements in \cite{croman-scaling-blockchain} were done,
the average block size was 540 KB.

\paragraph{Latency} The latency is the time required for a transaction obtain a
single confirmation (therefore the time required for the transaction to be
included in a block). It is roughly 10 minutes.





\subsubsection{Metric definitions}
\begin{definition}
  We define the metric ``X\% block propagation delay'' as the time required for
  X\% of the nodes to receive a full block
\end{definition}
\begin{remark}
  The calculation of this metric is done empirically, carrying out measurements
  as done in the experiment conducted in \cite{croman-scaling-blockchain}.
\end{remark}

\begin{definition}\label{def:effective-throughput}
  We define the metric ``X\% effective throughput'' as the block size divided by
  the metric ``X\% block propagation delay'':

  \[\text{X\% effective throughput} = \frac{\text{block size}}{\text{X\% block
  propagation delay}}\]
\end{definition}

\begin{remark}\label{remark-1}
  If the transactions rate (or, equivalently, the average block size) exceeds
  the X\% effective throughput during a block interval, then $(100-X)\%$
  of the nodes on the network would be unable to receive blocks as they arrive,
  so they would be effectively disabled. Table \ref{tab:effective-throughput-values}
  shows the connection between transactions rate and X\% effective throughput.
\end{remark}

\begin{remark}
  If the size or properties of the Bitcoin network change then the X\% effective
  throughput also changes.
\end{remark}

\begin{table}[h!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabularx}{\textwidth}{c c c}
\toprule
\textbf{X\%} & \textbf{Effective throughput} & \textbf{Transactions/sec}                                    \\ \midrule
50\% & 492 Kbps & 248 trans./sec \\
\\
90\% & 55 Kbps & 26 trans./sec \\ \bottomrule
\end{tabularx}
}
\caption{Effective throughputs and associated transactions per second. The values
are calculated assuming that 50\% and 90\% block propagation times are 8.7 seconds
and 79 seconds respectively, while the average block size is 540KB and the
transactions are 250-byte transactions. Note that the effective throughput
values are expressed in kilo-bits per second.}
\label{tab:effective-throughput-values}
\end{table}




%Interesting point offerend by article ***: their conclusion is that fundamental
%protocol redesign is needed for blockchains to scale significantly while
%retaining their decentralization (reparametrization only is not enough, as
%explained in SECTION 3). They also discuss about new strategies for designing
%new protocols by addressing blockchain limitations through a partition of the
%system in different layers, analyzing the bottlenecks and the limitations of
%each layer (SECTION 4).

\subsubsection{Throughput and latency limits}

\paragraph{Premises} It is assumed that it is desired to maintain nearly the
current level of decentralization in the system, namely it is desired to have at
least 90\% of properly functioning nodes. It is also assumed that the current
Bitcoin peer-to-peer network is used, without bringing any changes to it since
they would affect the X\% effective throughput.

\paragraph{Throughput limit} Given the two premises above, given the definition
\ref{def:effective-throughput} and given the remark \ref{remark-1}, the block size and
interval must satisfy:
\[ \frac{\text{block size}}{\text{X\% effective throughput}} < \text{block interval} \]
\emph{Consequence: given the current overlay network and today’s 10 min average
block interval, the block size should not exceed 4 MB \cite{croman-scaling-blockchain}.
A 4 MB block size corresponds to a throughput of at most 27 transactions/sec
($\text{maximum throughput} = \text{block size} / \text{block interval}$)}.


\paragraph{Latency limit} For improving the system latency a simple idea is to
reduce the block interval. In order to do so and maintain an high effective
throughput is however necessary to reduce the block size (consequence of the
throughput limit formula shown in the previous paragraph), but using a block too
small would not make full use of the network bandwidth. The following guideline
is the result of the experiments and measurements carried out in
\cite{croman-scaling-blockchain}:  \emph{given today’s overlay network, to retain at
least 90\% effective throughput and fully utilize the bandwidth of the network,
the block interval should not be significantly smaller than 12s.}





\subsection{Bitcoin bottlenecks analysis}  Besides the throughput and latency
limits discussed in the previous section, the Bitcoin system has some other
bottlenecks and limitations that will be  discussed in this section. In
particular, the system will be decomposed into a set of abstraction layers
called \emph{planes} (network, consensus, storage, view and side planes), which
allows limitations and bottlenecks to be addressed at each plane individually
and in a structured manner. Some solutions to this limitations will be discussed
as well.

\subsubsection{Network plane}
The function of the network plane is to propagate messages that represent
valid transactions.

Measurements done in \cite{croman-scaling-blockchain} showed that Bitcoin’s
network protocol does not fully utilize the available network bandwidth,
making Bitcoin’s Network Plane a bottleneck in transaction processing. The two
main inefficiencies of the plane are the following:
\begin{enumerate}
  \item In order to avoid denial-of-service caused by the propagation of invalid
  transactions, each node must fully receive and validate a transaction before
  further propagating it. This local validation of transactions done by each node
  contributes significantly to the overall propagation time.
  \item Bitcoin’s network protocol first propagates all transactions (in order
  to allow the nodes to create pools from which choose the transactions to include
  in the block they will mine) and then propagates a block (containing a set of
  the previously propagated transactions) again when it is mined. This causes
  some of the transactions to be transmitted twice.
\end{enumerate}

In order to remedy to the first inefficiency, a possible solution is proposed in
\cite{andresen}. The basic idea is to adopt a set reconciliation protocol in
which nodes only fetch transactions they do not possess in a newly mined block.
If in fact memory pools were perfectly synchronized, the miners could just
announce new blocks as the fixed 80-byte block header and the coinbase
transaction, without including all the transactions of the block. Memory pools
however are not perfectly synchronized but they're just similar: with a set
reconciliation protocol is possible to take advantage of this similarity for
optimizing the blocks broadcast\footnote{Set reconciliation means finding the
difference between two sets of data. In this case the sets are the set of
transactions transactions in a new block and the set of transactions that a node
already knows about (its memory pool).}.


\subsubsection{Consensus plane} This layer is responsible for mining and
achieving consensus. Bottlenecks in this plane are due to the proof of work
algorithms that introduce a tradeoff among consensus speed, bandwidth and
security since increasing the former two results in compromising security.
A solution to this limitation could be using different consensus algorithms, such
as the Proof of Stake discussed in section \ref{sec:proof-of-stake}.


\subsubsection{Storage plane} This plane functions as a global memory which
stores the Bitcoin ledger and supports the operations \code{read} and
\code{write}. The main issue with this layer is that each node has to store an
entire copy of the whole blockchain which leads to inefficiencies such as
increased bandwidth and storage requirements. The Bitcoin blockchain size, in
fact, is increasing exponentially and its size in June 2018 was around 178 GB
\cite{statista}.


\subsubsection{View plane} This plane proposes an optimization based on the fact
that miners for operating do not need the full blockchain that stores the entire
transaction history. Miners in Bitcoin in fact do not operate on the full
blockchain but, instead, they operate on a \emph{view} of the ledger called the
unspent transaction outputs set (UTXO set, see section \ref{}). A \emph{view}
can be considered as a data structure derived from the full ledger whose state
is obtained by applying all transactions.

The main issue is that the in Bitcoin the view is not stored in the Storage
plane and distributed to the nodes but instead the miners have to locally
compute it from the whole ledger from the beginning of time and this operation
now requires around four days, as well as the whole blockchain. Storing the
view in the storage plane instead would therefore eliminate the need for
mining nodes to store the full blockchain and improve efficiency.


\subsubsection{Side plane} This plane represents the off-chain transactions,
namely transactions that are not carried out over the blockchain but over an
out-of-band connection over the internet. These transactions are basically
transfer agreements which generally occur between two parties and are not stored
in the Blockchain, that, however, is still used to secure them. The advantages
of this transactions are the higher throughput (transactions are recorded
immediately without having to wait for network confirmations), the higher
privacy (transfers are not visible on the public blockchain) and the
inexpensiveness (there are no fees since there is no participant required to
validate the transaction). The main issue with these transactions is that they
rely critically on the nature of the links between the involved parties which
essentially form a separate network plane: reference
\cite{croman-scaling-blockchain} pointed out that this network can act as a
bottleneck as well.
